Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-04-26 numerical-particle.tex
edited n-body dynamics section

\enzo\ uses a particle-mesh N-body method to calculate the 
dynamics of collisionless systems \citep{Hockney88}.  This method 
follows trajectories of a representative sample of individual particles 
and is much more efficient than a direct solution of the Boltzmann 
equation in essentially all astrophysical situations. 
As described earlier, the gravitational potential is computed by solving the elliptic 
Poisson's equation (Eq.~\ref{eq:potential}), and differencing the potential
to find accelerations, which are then interpolated back to particles.  This
acceleration is time-centered (because the underlying gravitating mass
field is approximately time-centered), and so we have accelerations $\myvec{a}^{n+1/2}$
for each particle.  These are used to update the particle positions and velocities
starting from $\myvec{x}^n$ and $\myvec{v}^n$ using a standard drift-kick-drift technique:
\begin{eqnarray}
\label{eqn.driftkick}
\myvec{x}^{n+1/2} & = & \myvec{x}^n + \frac{\Delta t}{2 a^n} \myvec{v}^{n} \nonumber \\
\myvec{v}^{n+1} & = & \myvec{v}^n \left(1 - \frac{\dot{a}^{n+1/2}}{a^{n+1/2}}\right) + \frac{\Delta t}{ a^{n+1/2}} \myvec{a}^{n+1/2} \\
\myvec{x}^{n+1} & = & \myvec{x}^{n+1/2} + \frac{\Delta t}{2 a^{n+1}} \myvec{v}^{n+1} \nonumber
\end{eqnarray}

Particles are stored in the most highly refined grid patch at the point in space where they exist, and particles which move out of a subgrid patch are sent to the grid patch covering the adjacent volume with the finest spatial resolution, which may be of the same spatial resolution, coarser, or finer than the grid patch that the particles are moved from.  This takes place in a communication process at the end of each timestep on a level.

>>>>

\enzo\ uses a particle-mesh N-body method to calculate the dynamics of
collisionless systems \citep{Hockney88}.  This method follows
trajectories of a representative sample of individual particles that
sample the phase space of the dark matter distribution, and is much
more efficient than a direct solution of the Boltzmann equation in
essentially all astrophysical situations for the levels of accuracy
that are required for simulations of structure formation.  As
described earlier, the gravitational potential is computed by solving
the elliptic Poisson's equation (Eq.~\ref{eq:potential}), and
differencing the potential to find accelerations, which are then
interpolated back to particles.  This acceleration is time-centered
(because the underlying gravitating mass field is approximately
time-centered), and so we have accelerations $\myvec{a}^{n+1/2}$ for
each particle.  These are used to update the particle positions and
velocities starting from $\myvec{x}^n$ and $\myvec{v}^n$ using a
standard drift-kick-drift technique:

\begin{eqnarray}
\label{eqn.driftkick}
\myvec{x}^{n+1/2} & = & \myvec{x}^n + \frac{\Delta t}{2 a^n} \myvec{v}^{n} \nonumber \\
\myvec{v}^{n+1} & = & \myvec{v}^n \left(1 - \frac{\dot{a}^{n+1/2}}{a^{n+1/2}}\right) + \frac{\Delta t}{ a^{n+1/2}} \myvec{a}^{n+1/2} \\
\myvec{x}^{n+1} & = & \myvec{x}^{n+1/2} + \frac{\Delta t}{2 a^{n+1}} \myvec{v}^{n+1} \nonumber
\end{eqnarray}

Particles are stored in the most highly refined grid patch at the
point in space where they exist, and particles that move out of a
subgrid patch are sent to the grid patch covering the adjacent volume
with the finest spatial resolution, which may be of the same spatial
resolution, coarser, or finer than the grid patch that the particles
are moved from.  This takes place in a communication process at the
end of each timestep on a level.
